---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
%https://github.com/patbr21/RepData_PeerAssessment1/upload/master
##Used librarys is this assignment
```r{}
library(readr)
library(dplyr)
```
## Loading and preprocessing the data
First we load the data via readr. It is stored in a variable called 'activity'.
We use the library readr for this task.
```{r, cache=TRUE}
activity <- read_csv("activity.csv")
```
 Let's have a look at the first 10 lines after reading the dataset.
 ```{r}
head(activity)
```
## What is mean total number of steps taken per day?
First we calculate the number of steps per day, then the resulting mean. We are using the dplyr package for the calculation.
```{r}
activity%>%
          group_by(date)%>%
          summarise(steps_day = sum(steps, na.rm = TRUE))
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
